\documentclass[A4 paper]{article}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage[english]{babel}
\usepackage{amsmath, amsfonts, graphicx, float}
\usepackage{pgf,tikz}
\usepackage{pgfpages}
\usetikzlibrary{arrows}
\DeclareMathOperator{\Mod}{mod}

\title{Problem Set 10} 
\author{Written by James Bang for the NSW AMOC Correspondence Program}
\date{21th December 2019}
\begin{document} 
\maketitle
\thispagestyle{empty}
\section*{Instructions}
\begin{itemize}
\item This problem set is based off the notes ``\textit{Polynomials}". 
\item They are in roughly difficulty order and get quite difficult, so you \textbf{not} expected to be able to solve every problem. 
\item However, please attempt as many questions as you can and submit your solutions to your mentor for marking and feedback.
\item You may type your solutions or take a \textbf{clear} scan/photo of \textbf{legible} written solutions.
\item Feel free to discuss these problems with your peers and on the forum but the solutions you submit must be written by yourself.
\end{itemize}
\section*{Problems}
\begin{enumerate}
	\item Let $a,b,c$ be the roots of the polynomial $P(x)=2x^3-3x^2+4x-5$. Evaluate the following numbers in closed form (i.e. express them as a single rational number).
	\begin{enumerate}
		\item $a+b+c$;
		\item $a^2+b^2+c^2$;
		\item $(a+b)(b+c)(c+a)$.
	\end{enumerate}
	\item Let $x\neq y$ be integers. Show, by factorisation or otherwise, that $x-y\mid x^n-y^n$. Hence show that $x-y\mid P(x)-P(y)$ whenever $P(x)$ is an integer polynomial.
	\item Let $P(x)$ be a \textbf{cubic} polynomials with roots $r_1,r_2,r_3$. It is given that \[\frac{P(\frac{1}{2})+P(-\frac{1}{2})}{P(0)}=1000.\] Find the value of $\frac{1}{r_1r_2}+\frac{1}{r_2r_3}+\frac{1}{r_3r_1}$.
	
	
\item A nonzero real number $a$ is given. Find all polynomials $P(x)$ with real coefficients for which $P(x+a)=P(x)+a$ for every real number $x$.

\vspace{2mm}

\noindent \textit{Hint. What can you say about the polynomial $Q(x)=P(x)-x$?}


\item Find all polynomials $P(x)$ with real coefficients such that \[(x-27)P(3x)=27(x-1)P(x).\]

\end{enumerate}
 
\end{document}
